---
title: "Is there a difference between Nigerian an Ghananian afrobeats songs?"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

```{r, setup}
library(tidyverse)
library(plotly)
library(spotifyr)
library(compmus)
library(flexdashboard)
```

### Introduction

Afrobeats is a music genre originated from West Africa. More specifically from Nigeria and Ghana. It was very popular in the 2000's and 2010's. The artists sing mostly in Nigerian or Ghanaian accented English, or in the languages originated from these countries. Afrobeats is also called afropop or afrofusion. It's a fusion of different genres like dancehall, R&B, hip hop, etc. I think this genre is very interesting because the songs have an unique sound, and reflect the culture and history of West Africa. It's fairly new that people from all over the world listen to this music.

Other than the languages spoken in the song I'm interested to look for other differences between Nigerian and Ghanaian songs. I want to look for the unique influences and cultural heritage of each country in the different songs. Popular Nigerian artists are Wizkid, Tems and Burna Boy. Stonebwoy and Gyakie are examples of Ghanian singers. I'm going to use playlists with afrobeats music originated from both these countries. These playlists covers the Afrobeats genre well, but there are probably songs in it from artists who aren't originally from Nigeria or Ghana.

'Higher' from Tems, a Nigerian singer, is a slow song, which seems to sound like 'SOMETHING' from Gyakie, a Ghanian singer. These songs from these two women, are very typical for the afrobeats genre. It's interesting to compare these two songs and look for differences and similarities.

### The plot

The plot I made is comparing the tempo and danceability of songs on the x and y axis. Also it's showing the energy of songs with colour, and the popularity with size. It's interesting to see how Ghanian songs are more scattered over the x axis, which means the tempo is different for every song. The tempo of Nigerian songs are more similar to eachother. It's also interesting to look if tempo and dancebility are depended variables, because intuitively it is.


```{r}
ghana <- get_playlist_audio_features("","37i9dQZF1DX3zd1ZTQB0K9?si=71424602f1f446bf")
nigeria <- get_playlist_audio_features("", "5oLoDUUcDfJPfMXuwXMiXb?si=b5454940a9914127")

ghana_nigeria <-
  bind_rows(
    ghana |> mutate(category = "Ghana"),
    nigeria |> mutate(category = "Nigeria")
  )

ghana_nigeria |>                   
  ggplot(aes(x = tempo, y = danceability, size = track.popularity, colour = liveness)) + 
  geom_point() + 
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
  geom_text(aes(x = tempo, y = danceability, label = label),
  data = 
    tibble(label = c("O2", "Asiwaju"),category = c("Ghana", "Nigeria"),tempo = c(66.769, 199.796),danceability = c(0.437, 0.727)),
  colour = "black",size = 3,hjust = "right", nudge_y = -0.02) +       
  facet_wrap(~ category) + 
  scale_x_continuous(limits = c(50, 200), breaks = c(50, 100, 200), minor_breaks = NULL) +
  scale_y_continuous(limits = c(0.4, 1),breaks = c(0.4, 0.7, 1),minor_breaks = NULL) +
  scale_size_continuous(guide = "none") +
  theme_light() + labs(x = "Tempo", y = "Dancability",colour = "Energy")

```
### Conclusion and discussion


